---
title: "Fixed effect BLUEs"
author: "Harly Durbin"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyr)
library(stringr)
library(dplyr)
library(glue)
library(ggplot2)
library(magrittr)
library(tidylog)

source(here::here("source_functions/melt_aireml.R"))
source(here::here("source_functions/parse_renf90table.R"))

```

# Notes & questions


```{r}
breed_key <- read_rds(here::here("data/derived_data/breed_key/breed_key.rds"))
```

# Setup 

```{r, warning=FALSE, message=FALSE}
cleaned <- read_rds(here::here("data/derived_data/import_join_clean/cleaned.rds"))
```

```{r, warning=FALSE, message=FALSE}
full_ped <- read_rds(here::here("data/derived_data/3gen/full_ped.rds"))
```

## Age model

```{r, warning=FALSE, message=FALSE}
fixed16_sol <-
  c(1, 2) %>% 
  purrr::map_dfr(~ parse_renf90table(here::here("data/derived_data/aireml_varcomp/fixed16/renf90.tables"),
                                     effect_num = .x,
                                     effect_key = TRUE)) %>% 
  mutate(effect = case_when(effect == 1 ~ "cg",
                            effect == 2 ~ "age_group")) %>% 
  right_join(read_table2(here::here("data/derived_data/aireml_varcomp/fixed16/solutions"),
                         skip = 1,
                         col_names = c("trait", "effect", "id_renamed", "solution", "se")) %>%
               select(-trait) %>% 
               filter(!effect %in% c(3, 4)) %>% 
               mutate_at(vars("effect", "id_renamed"), ~ as.character(.)) %>% 
               mutate(effect = case_when(effect == 1 ~ "cg",
                                         effect == 2 ~ "age_group")))
  
  
```

# Age

Supplementary Figure XXX. Best linear unbiased estimates for age in years fit as a categorical effect were used to determine age grouping in contemporary group definition. Records from very old animals aged 14-21 were combined to increase sample size. 

```{r}
fixed16_sol %>% 
  filter(effect == "age_group") %>% 
  mutate(id_original = if_else(id_original == "old", 14, as.numeric(id_original))) %>% 
  arrange(id_original) %>% 
  mutate(id_original = forcats::fct_inorder(as.factor(id_original)),
         rel_sol = solution + 0.20198118,
         clr = case_when(rel_sol > 0 ~"#0070C0",
                         TRUE ~ "lightgrey")) %>% 
  ggplot(aes(x = id_original,
             #y = solution,
             y = rel_sol,
             fill = clr)) +
  geom_bar(stat = "identity",
           position = "identity") +
  scale_fill_identity() +
  geom_errorbar(aes(ymin = rel_sol - se,
                    ymax = rel_sol + se)) +
  theme_classic() +
  scale_x_discrete(labels = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14+")) +
  theme(plot.title = element_text(size = 22,
                                  face = "italic",
                                  margin = margin(t = 0,
                                                  r = 0,
                                                  b = 13,
                                                  l = 0)),
        axis.title = element_text(size = 22),
        axis.title.y = element_text(margin = margin(t = 0,
                                                    r = 10,
                                                    b = 0,
                                                    l = 0)),
        axis.title.x = element_text(margin = margin(t = 10,
                                                    r = 0,
                                                    b = 0,
                                                    l = 0)),
        axis.text = element_text(size = 16),
        legend.position = "none",
        panel.background = element_rect(fill = "transparent",
                                        colour = NA),
        plot.background = element_rect(fill = "transparent",
                                       colour = NA),
        legend.background = element_rect(fill = "transparent",
                                       colour = NA)) +
  labs(x = "Age in years",
       y = "Effect estimate")
```


```{r}
ggsave(here::here("figures/misc_blues/fixed_16.ages.transparent_relative.png"), width = 8, height = 5, bg = "transparent")
```

# Calving season, fescue, and age group BLUEs from environmental models 

## Temperature only (`fixed6`)

```{r, warning=FALSE, message=FALSE}
c(1:4) %>% 
  purrr::map_dfr(~ parse_renf90table(here::here("data/derived_data/aireml_varcomp/fixed6/renf90.tables"),
                                     effect_num = .x,
                                     effect_key = TRUE)) %>% 
  left_join(read_table2(here::here("data/derived_data/aireml_varcomp/fixed6/solutions"),
                        skip = 1,
                        col_names = c("trait", "effect", "id_renamed", "solution", "se")) %>%
              mutate(id_renamed = as.character(id_renamed))) %>% 
  filter(effect %in% c(2, 3, 4))
```

## Day length only (`fixed5`)

```{r, warning=FALSE, message=FALSE}

c(1:4) %>% 
  purrr::map_dfr(~ parse_renf90table(here::here("data/derived_data/aireml_varcomp/fixed5/renf90.tables"),
                                     effect_num = .x,
                                     effect_key = TRUE)) %>% 
  left_join(read_table2(here::here("data/derived_data/aireml_varcomp/fixed5/solutions"),
                        skip = 1,
                        col_names = c("trait", "effect", "id_renamed", "solution", "se")) %>%
              mutate(id_renamed = as.character(id_renamed))) %>% 
  filter(effect %in% c(2, 3, 4))

```

## Temperature + day length (`fixed3`)

```{r, warning=FALSE, message=FALSE}
c(1:4) %>% 
  purrr::map_dfr(~ parse_renf90table(here::here("data/derived_data/aireml_varcomp/fixed3/renf90.tables"),
                                     effect_num = .x,
                                     effect_key = TRUE)) %>% 
  left_join(read_table2(here::here("data/derived_data/aireml_varcomp/fixed3/solutions"),
                        skip = 1,
                        col_names = c("trait", "effect", "id_renamed", "solution", "se")) %>%
              mutate(id_renamed = as.character(id_renamed))) %>% 
  filter(effect %in% c(2, 3, 4))
```

## Most complicated model: temperature + day length + centered temperature*day length (`fixed13`)

```{r, warning=FALSE, message=FALSE}
c(1:4) %>% 
  purrr::map_dfr(~ parse_renf90table(here::here("data/derived_data/aireml_varcomp/fixed13/renf90.tables"),
                                     effect_num = .x,
                                     effect_key = TRUE)) %>% 
  left_join(read_table2(here::here("data/derived_data/aireml_varcomp/fixed13/solutions"),
                        skip = 1,
                        col_names = c("trait", "effect", "id_renamed", "solution", "se")) %>%
              mutate(id_renamed = as.character(id_renamed))) %>% 
  filter(effect %in% c(2, 3, 4))
```
